---
title: "Exploratory markdown"
author: "Hannah Haskell"
date: "4/15/2021"
output: pdf_document
---
```{r setup, include=FALSE}
#setting the working directory
knitr::opts_knit$set(root.dir = '/Users/hannahhaskell/Desktop/')
```

## Libraries

`tidyverse`

`lubridate`

`gapminder`

`modelr`

`broom`

## Data

Two data files will be used for this exploratory analysis. The CEA_data includes  physiological measurements of the oysters during the exposure trials as well as size measurements, mortalities, and upper lethal temperatures. The algae data includes the number of algae cells cleared over time during a post exposure clearance rate trial.  

```{r loading the data, include=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(gapminder)
library(knitr)
library(modelr)
library(broom)

CEA_data <- read_csv("Project_Mock_CEA_data.csv")
Algae <- read_csv("Project_Mock_Algae_clearing.csv")

```

## Basic Data Exploration

### CEA Data:

**Rows:** n = `r nrow(CEA_data)`

**Columns:** n = `r ncol(CEA_data)`

**Column Names:** `r names(CEA_data)`

### Algae Data:

**Rows:** n = `r nrow(Algae)`

**Columns:** n = `r ncol(Algae)`

**Column Names:** `r names(Algae)`

After looking over the column names the CEA data seems a bit disorganized. It looks like there was some spaces left between columns that should be removed and I think it is worth while to assign a few more variables to help separate and organize all of the data. 
